---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(ggplot2)
data <- read.csv('activity.csv')
data$interval <- sprintf('%02d:%02d',data$interval%/%100, data$interval%%100)
data$time <- strptime(with(data, paste(date, interval)), '%Y-%m-%d %H:%M')
```


## What is mean total number of steps taken per day?
```{r}
d <- data[!is.na(data$steps), ]
steps_perday <- sapply(split(d,d$date), function(x) sum(x$steps))
#hist(steps_perday)
qplot(steps_perday, binwidth = 2500)
```

```{r}
mean(steps_perday)
median(steps_perday)
```


## What is the average daily activity pattern?

```{r}
steps <- sapply(split(d, d$interval), function(x) mean(x$step))
df <- data.frame(steps = steps, time = names(steps))
plot(strptime(df$time, format = '%H:%M'), df$steps, type = 'l', xlab = 'time', ylab = 'steps')
#plot(strptime(names(steps), format='%H:%M'), steps, type='l', xlab='time')
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
names(which.max(steps))
df$time[which.max(df$steps)]
```

## Imputing missing values
Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(data$steps))
```

Create a new dataset that is equal to the original dataset but with the missing data filled in, using the mean for that 5-minute interval
```{r}
nasteps <- is.na(data$steps)
data_new <- data
data_new$steps[nasteps] <- steps[data_new$interval[nasteps]]
```

```{r}
steps_perday_new <- sapply(split(data_new,data_new$date), function(x) sum(x$steps))
hist(steps_perday_new)
mean(steps_perday_new)
median(steps_perday_new)
```
These values are greater than the estimates from the first part, 

## Are there differences in activity patterns between weekdays and weekends?
```{r}
data_new$weekday <- weekdays(data_new$time) == 'Sunday' | weekdays(data_new$time) == 'Saturday'
data_new$weekday <- factor(data_new$weekday, label=c('Weekday', 'Weekend'))
d_weekend <- data_new[data_new$weekday=='Weekend',]
d_weekday<- data_new[data_new$weekday=='Weekday',]
steps_weekend <- sapply(split(d_weekend, d_weekend$interval), function(x) mean(x$step))
df_weekend <- data.frame(steps = steps_weekend, time = names(steps_weekend), Weekday = 0)
steps_weekday <- sapply(split(d_weekday, d_weekday$interval), function(x) mean(x$step))
df_weekday <- data.frame(steps = steps_weekday, time = names(steps_weekday), Weekday = 1)

df_new <- rbind(df_weekday, df_weekend)
df_new$Weekday <- factor(df_new$Weekday, labels = c('weekend', 'weekday'))
qplot(strptime(time, format = '%H:%M'), steps, data = df_new, facets = Weekday ~ ., geom = 'line', xlab = 'time')

```